Language-Based Security for Low-Level MPC

### 2024-07-16 ppdp24/model.tex
checkpoint

We define discrete joint probability mass functions (pmfs) in a
standard manner but develop some notations that are useful for our
presentation. Firstly, whereas distributions typically map
realizations of random variables to values in $[0..1]$, for
convenience and following \cite{barthe2019probabilistic} we use
memories to represent realizations, so for example given a pmf $\pmf$
over variables $\{ \sx{x}{1}, \mx{y}{2} \}$ we write $\pmf(\{
\elab{\secret{x}}{1} \mapsto 0, \elab{\mesg{y}}{2} \mapsto 1 \})$ to
denote the (joint) probability that $\elab{\secret{x}}{1} = 0 \wedge
\elab{\mesg{y}}{2} = 1$. Recall from Section
\ref{section-lang-semantics} that $\uplus$ denotes the combination of
memories, so for example $\{ \elab{\secret{x}}{1} \mapsto 0\} \uplus
\{\elab{\mesg{y}}{2} \mapsto 1 \} = \{ \elab{\secret{x}}{1} \mapsto 0,
\elab{\mesg{y}}{2} \mapsto 1 \}$.
\begin{definition}
  A \emph{probability mass function} $\pmf$ is a function
  mapping memories in $\mems(X)$ for given variables $X$ to
  values in $\mathbb{R}$ such that:
  $$
  \sum_{\store \in \mems(X)} \pmf(\store) \  = \ 1
  $$
  %The \emph{support} of a distribution is the set of realizations
  %of random variables with non-zero probability:
  %$\support(\pmf) \defeq \{ (v_1,\ldots,v_n) \mid
  %\pmf(\{ x_1 \mapsto v_1, \ldots, x_n \mapsto v_n\}) > 0 \} $
\end{definition}
%To recover succinct and familiar notation, we may omit the domain of a
%distribution when it is clear from an application context-
%i.e., we allow the following sugaring:
%$$
%\pdf{}(\store) \defeq \pdf{\dom(\store)}(\store)
%$$
Now, we can define a notion of marginal and conditional
distributions as follows, which are standard for discrete
probability mass functions. 
\begin{definition}
  Given $\pmf$ with $\dom(\pmf) = \mems(X_2)$, the \emph{marginal distribution}
  of variables $X_1 \subseteq X_2$ in $\pmf$, denoted $\margd{\pmf}{X_1}$, is defined as follows:
  $$
  \forall \store \in \mems(X_1) \quad . \quad \margd{\pmf}{X_1}(\store) =
  \sum_{\store' \in \mems(X_2-X_1)} \pmf(\store \uplus \store')
  $$
\end{definition}

\begin{definition}
  Given $\pmf$ with $\dom(\pmf) = \mems(X_3)$, the \emph{conditional distribution}
  of $X_1$ given $X_2$ where $X_1 \cup X_2 \subseteq X_3$ and $X_1 \cap X_2 = \varnothing$, denoted $\condd{\pmf}{X_1}{X_2}$,
  is defined as follows:
  $$
  \forall \store \in \mems(X_1 \cup X_2)\ .\ 
  \condd{\pmf}{X_1}{X_2}(\store) =
  \begin{cases}
    0 \text{\ if\ } \margd{\pmf}{X_2}(\store_X_2) = 0\\
    \margd{\pmf}{X_1 \cup X_2}(\store) / \margd{\pmf}{X_2}(\store_X_2)\ \text{\ otherwise}
  \end{cases}
  $$
\end{definition}
We also define some convenient syntactic sugarings. The first will allow us to
compare marginal distributions under different realization conditions
(as in, e.g., Definition \ref{definition-NIMO}), the others are standard.
\begin{definition}
  Given $\pmf$, for all $\store_2 \in \mems(X_2)$ and $\store_1 \in \mems(X_1)$ define:
  \begin{enumerate}
  \item $\condd{\pmf}{X_1}{\store_2}(\store_1) \defeq \condd{\pmf}{X_1}{X_2}(\store_1 \uplus \store_2)$
  \item $\pmf(\store_1)  \defeq \margd{\pmf}{X_1}(\store_1)$ 
  \item $\pmf(\store_1|\store_2) \defeq \condd{\pmf}{X_1}{X_2}(\store_1 \uplus \store_2)$
  \end{enumerate}
\end{definition}

>>>>

We define discrete joint probability mass functions (pmfs) in a
standard manner but develop some notations that are useful for our
presentation. Firstly, whereas distributions typically map
realizations of random variables to values in $[0..1]$, for
convenience and following \cite{barthe2019probabilistic} we use
memories to represent realizations, so for example given a pmf $\pmf$
over variables $\{ \sx{x}{1}, \mx{y}{2} \}$ we write $\pmf(\{
\elab{\secret{x}}{1} \mapsto 0, \elab{\mesg{y}}{2} \mapsto 1 \})$ to
denote the (joint) probability that $\elab{\secret{x}}{1} = 0 \wedge
\elab{\mesg{y}}{2} = 1$. Recall from Section
\ref{section-lang-semantics} that $\uplus$ denotes the combination of
memories, so for example $\{ \elab{\secret{x}}{1} \mapsto 0\} \uplus
\{\elab{\mesg{y}}{2} \mapsto 1 \} = \{ \elab{\secret{x}}{1} \mapsto 0,
\elab{\mesg{y}}{2} \mapsto 1 \}$.
\begin{definition}
  A \emph{probability mass function} $\pmf$ is a function
  mapping memories in $\mems(X)$ for given variables $X$ to
  values in $\mathbb{R}$ such that:
  $$
  \sum_{\store \in \mems(X)} \pmf(\store) \  = \ 1
  $$
  %The \emph{support} of a distribution is the set of realizations
  %of random variables with non-zero probability:
  %$\support(\pmf) \defeq \{ (v_1,\ldots,v_n) \mid
  %\pmf(\{ x_1 \mapsto v_1, \ldots, x_n \mapsto v_n\}) > 0 \} $
\end{definition}
%To recover succinct and familiar notation, we may omit the domain of a
%distribution when it is clear from an application context-
%i.e., we allow the following sugaring:
%$$
%\pdf{}(\store) \defeq \pdf{\dom(\store)}(\store)
%$$
Now, we can define a notion of marginal and conditional
distributions as follows, which are standard for discrete
probability mass functions. 
\begin{definition}
  Given $\pmf$ with $\dom(\pmf) = \mems(X_2)$, the \emph{marginal distribution}
  of variables $X_1 \subseteq X_2$ in $\pmf$ is denoted $\margd{\pmf}{X_1}$ and defined as follows:
  $$
  \forall \store \in \mems(X_1) \quad . \quad \margd{\pmf}{X_1}(\store) \defeq
  \sum_{\store' \in \mems(X_2-X_1)} \pmf(\store \uplus \store')
  $$
\end{definition}

\begin{definition}
  Given $\pmf$, the \emph{conditional distribution}
  of $X_1$ given $X_2$ where $X_1 \cup X_2 \subseteq \dom(\pmf)$ and $X_1 \cap X_2 = \varnothing$
  is denoted $\condd{\pmf}{X_1}{X_2}$ and defined as follows:
  $$
  \forall \store \in \mems(X_1 \cup X_2)\ .\ 
  \condd{\pmf}{X_1}{X_2}(\store) \defeq
  \begin{cases}
    0 \text{\ if\ } \margd{\pmf}{X_2}(\store_{X_2}) = 0\\
    \margd{\pmf}{X_1 \cup X_2}(\store) / \margd{\pmf}{X_2}(\store_{X_2})\ \text{\ o.w.}
  \end{cases}
  $$
\end{definition}
We also define some convenient syntactic sugarings. The first will allow us to
compare marginal distributions under different realization conditions
(as in, e.g., Definition \ref{definition-NIMO}), the others are standard.
\begin{definition}
  Given $\pmf$, for all $\store_1 \in \mems(X_1)$ and $\store_2 \in \mems(X_2)$ define:
  \begin{enumerate}
  \item $\condd{\pmf}{X_1}{\store_2}(\store_1) \defeq \condd{\pmf}{X_1}{X_2}(\store_1 \uplus \store_2)$
  \item $\pmf(\store_1)  \defeq \margd{\pmf}{X_1}(\store_1)$ 
  \item $\pmf(\store_1|\store_2) \defeq \condd{\pmf}{X_1}{X_2}(\store_1 \uplus \store_2)$
  \end{enumerate}
\end{definition}
